# RepData Peer Assesment 1
Markdown practice using activity data from a device such as a Fitbit.

## Data Source
The data contained in a zip file was retireved from https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip on March 13 2015.

### Required Libraries
```{r, echo = T, message=FALSE}
library(lubridate)
library(dplyr)
library(ggplot2)
library(lattice)
```

### Load the data
```{r, echo = T}
# Create a class that can convert from a date in a character format to a Date class
setClass("lDate")
setAs("character","lDate", function(from) ymd(from) )

# unzip the activity.csv into the current working directory
unzip("repdata-data-activity.zip")

# Read in the acitivyt.csv file
activity <- read.csv("activity.csv", 
                     stringsAsFactors = F, 
                     colClasses = c("numeric", "lDate", "numeric"), 
                     header = T)
```

### Summarize the data
```{r, echo = T}
summary(activity, na.rm=T)
```


### Mean steps taken per day

```{r, echo=T}
total_steps_per_day <- 
  activity %>% 
  group_by(date) %>% 
  summarise(total.steps = sum(steps, na.rm = T))
```

Histogram of total steps taken per day

```{r, echo=T}
ggplot(total_steps_per_day) +
  geom_histogram(aes(x=total.steps), bindwith = 61, colour="black", fill="blue") +
  ggtitle("Frequency of total steps per day")
```

Median and mean of the total steps taken per day

```{r, echo=T}
total_steps_per_day %>%
  summarise(median.steps = median(total.steps, na.rm = T), mean.steps = mean(total.steps, na.rm = T))
```


### Average activity pattern

```{r, echo=T}
mean_interval_activity <- 
  activity %>%
  group_by(interval) %>%
  summarise(mean.steps = mean(steps, na.rm = T))

ggplot(mean_interval_activity) +
  geom_line(aes(x = interval, y = mean.steps), size = 1, color = "blue") +
  ggtitle("Average steps for each interval for all days")

# Calculate interval with max average number of steps
mean_interval_activity[which.max(mean_interval_activity$mean.steps),]
```

### Missing Data Analysis
Total number of missing values

```{r, echo=TRUE}
sapply(activity, function(x) sum(is.na(x)))
```


Replace NA step values with mean step value for that interval
```{r, echo=TRUE}
pactivity <- activity
for(idx in which(is.na(pactivity$steps), arr.ind = T)){
  pactivity[idx, 1] <- mean_interval_activity[mean_interval_activity$interval == pactivity[idx, 3], 2]
}
```

Mean steps taken per day

```{r, echo=T}
total_steps_per_day2 <- 
  pactivity %>% 
  group_by(date) %>% 
  summarise(total.steps = sum(steps, na.rm = T))
```

Histogram of total steps taken per day

```{r, echo=T}
ggplot(total_steps_per_day2) +
  geom_histogram(aes(x=total.steps), bindwith = 61, colour="black", fill="blue") +
  ggtitle("Frequency of total steps per day")
```

Median and mean of the total steps taken per day

```{r, echo=T}
total_steps_per_day2 %>%
  summarise(median.steps = median(total.steps, na.rm = T), mean.steps = mean(total.steps, na.rm = T))
```
  
High concentrations of steps incrased, low concentraions decreased.

The median incrased slightly and the median equals the mean.

By using the mean of the interval to fill NA values the distribution of steps changed.

### Weekday and Weekend Activity

```{r, echo=TRUE}
# Make a new variable for each observation that says if the observation is on a weekday or weekend
pactivity$week.type <- rep('weekday', nrow(pactivity))
pactivity$week.type[grep('Sat|Sun', weekdays(pactivity$date))] <- "weekend"
pactivity$week.type <- as.factor(pactivity$week.type)
```

#### Plot average number of steps taken, averaged across all weekday days or weekend days

```{r, echo=T}
avg_activity_by_weektype <-
  pactivity %>%
  group_by(interval, week.type) %>%
  summarize(mean.steps = mean(steps))

ggplot(avg_activity_by_weektype, aes(color = week.type)) +
  geom_line(aes(interval, mean.steps), size = 1) +
  facet_grid(. ~ week.type) +
  theme(legend.position = 'none') +
  labs(title = "Average Steps per 5-minute Inverval for Weekdays and Weekends")
```

